---
title: "Course4_wk1_project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the data

#### When loading the dataset into R, please consider the following:
* The dataset has 2,075,259 rows and 9 columns. First calculate a rough estimate of how much memory the dataset will require in memory before reading into R. Make sure your computer has enough memory (most modern computers should be fine).
```{r}
row <- 2075259
col <- 9
# Assume all are numeric data. on most modern computers double precision floating point numbers are stored using 64 bits of memory, or 8 bytes. Given that information, you can do the following calculation
memory <- (row * col * 8)/2^{20}
# In megabytes
memory
# In gigabytes
memory/1000
# Most computers these days have at least that much RAM.
```

* We will only be using data from the dates 2007-02-01 and 2007-02-02. One alternative is to read the data from just those dates rather than reading in the entire dataset and subsetting to those dates.


* You may find it useful to convert the Date and Time variables to Date/Time classes in R using the strptime() and as.Date() functions.

* Note that in this dataset missing values are coded as ?.

```{r}
setwd("D:/Online Courses/Data Science Foundations using R Specialization/Course 4 Exploratory Data Analysis/Week 1/exdata_data_household_power_consumption")
df = read.table('household_power_consumption.txt', sep=";", nrows = 300000, na.strings = "?", header=TRUE)
names(df)
tail(df,2)
str(df)

# convert the date and time variable to Date and time class
df$Date <- as.Date(df$Date, format = "%d/%m/%Y")
head(df)
df$Time <- format(strptime(df$Time, "%H:%M:%S"),"%H:%M:%S")

# Subset the data
df <- subset(df, subset = (Date >= "2007-02-01" & Date <= "2007-02-02"))
head(df)
```
#### Plot plot1
```{r}
hist(df$Global_active_power, main = "Global Active Power", 
     xlab = "Global Active Power (kilowatts)", col = "Red")
# Save plot1
dev.copy(png, file = "plot1.png", height = 480, width = 480)
dev.off()
```

#### Plot plot2
```{r}
# Convert dates and times
df$datetime <- strptime(paste(df$Date, df$Time), "%Y-%m-%d %H:%M:%S")
df$datetime <- as.POSIXct(df$datetime)
plot(df$Global_active_power ~ df$datetime, type = "l", 
     xlab = "", ylab = "Global Active Power (kilowatts)")
# Save plot2
dev.copy(png, file = "plot2.png", height = 480, width = 480)
dev.off()
```

#### Plot plot3
```{r}
x <-  df$datetime
y1 <- df$Sub_metering_1
y2 <- df$Sub_metering_2
y3 <- df$Sub_metering_3
plot(x, y1,type="l", ylab = "Energy sub metering", xlab = "")
lines(x,y2,col="red")
lines(x,y3, col ="blue")
legend("topright", lty = 1, col = c("black", "red", "blue"), legend = c("Sub_metering_1", "Sub_metering_2", "Sub_metering_3"))
# Save plot3
dev.copy(png, file = "plot3.png", height = 480, width = 480)
dev.off()
```
#### Plot plot4
```{r}
par(mfrow = c(2, 2))
plot(df$Global_active_power ~ df$datetime, type = "l", 
     xlab = "", ylab = "Global Active Power")

plot(df$Voltage ~ df$datetime, type = "l", xlab = "datetime", ylab = "Voltage")

plot(x, y1,type="l", ylab = "Energy sub metering", xlab = "")
lines(x,y2,col="red")
lines(x,y3, col ="blue")
legend("topright", lty = 1, bty = "n", col = c("black", "red", "blue"), legend = c("Sub_metering_1", "Sub_metering_2", "Sub_metering_3"))

plot(df$Global_reactive_power ~ df$datetime, type = "l", xlab = "datetime", ylab = "Global_reactive_power")

dev.copy(png, file = "plot4.png", height = 480, width = 480)
dev.off()
```
